---
title: "Heart Shaped Network Analysis File 02 - Final Visualisation"
author: "Developed by Honor McGregor with the patient assistance of ChatGPT"
date: "2026-07-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# NETWORK ANALYSIS CONSTRAINED TO A HEART

```{r}
library(tidyverse)
library(igraph)
library(tidygraph)
library(ggraph)
library(ggrepel)
```

## 01 Rebuild the full network safely

```{r}
edges_full <- edges10 |>
  select(from, to, weight)

nodes_full <- tibble(
  name = sort(unique(c(edges_full$from, edges_full$to)))
) |>
  left_join(
    node_counts |>
      transmute(
        name = word,
        Count = count
      ),
    by = "name"
  ) |>
  mutate(
    Count = if_else(is.na(Count), 1L, Count),
    Word = name
  )
```



**Now detect communities:**

```{r}
g_temp <- igraph::graph_from_data_frame(
  d = edges_full,
  vertices = nodes_full,
  directed = FALSE
)

communities_full <- igraph::cluster_louvain(
  g_temp,
  weights = igraph::E(g_temp)$weight
)

nodes_full <- nodes_full |>
  mutate(
    Group = as.character(igraph::membership(communities_full)[name])
  )
```

# ============================================================
# IMPORTANT: CHECK THE NUMBER OF COMMUNITIES
# ============================================================
#
# Community detection is performed again here for the network
# used in the heart visualisation.
#
# The number of communities may differ from those identified
# earlier in the text/network analysis (File 01 fb_heart_data.Rmd).
#
# Check the communities before continuing:

table(nodes_full$Group)

# Note the number of communities returned.
# You will need this number when setting up:
#   - community anchors
#   - community labels
#   - community colours
#   - community density layers

**Now rebuild the graph with the community information included:**

```{r}
g_full <- igraph::graph_from_data_frame(
  d = edges_full,
  vertices = nodes_full,
  directed = FALSE
)

graph_full <- as_tbl_graph(g_full)

layout_base_full <- create_layout(
  graph_full,
  layout = "stress"
)

nrow(layout_base_full)
```

## 2. Create the mathematical heart properly

```{r}
t <- seq(0, 2*pi, length.out = 50000)

heart_raw <- tibble(
  t = t,
  x = 16 * sin(t)^3,
  y = 13*cos(t) - 5*cos(2*t) - 2*cos(3*t) - cos(4*t)
) |>
  mutate(
    x = x - mean(x),
    y = y - mean(y),
    theta = atan2(y, x),
    r = sqrt(x^2 + y^2)
  )

max_heart_r <- max(heart_raw$r)

heart_boundary <- heart_raw |>
  transmute(
    x = x / max_heart_r,
    y = y / max_heart_r
  )

heart_lookup <- heart_raw |>
  mutate(
    angle_bin = round(theta, 3)
  ) |>
  group_by(angle_bin) |>
  summarise(
    heart_radius = max(r) / max_heart_r,
    .groups = "drop"
  ) |>
  arrange(angle_bin)

R_heart <- approxfun(
  x = heart_lookup$angle_bin,
  y = heart_lookup$heart_radius,
  rule = 2
)
```

## 3. Map the full network into the heart

```{r}
layout_heartmap_full <- layout_base_full |>
  as_tibble() |>
  mutate(
    x_centre = (min(x) + max(x)) / 2,
    y_centre = (min(y) + max(y)) / 2,

    x0 = x - x_centre,
    y0 = y - y_centre,

    theta = atan2(y0, x0),
    r = sqrt(x0^2 + y0^2),

    r_norm = r / max(r),

    # smaller exponent = more spread toward the heart boundary
    r_expanded = r_norm^0.30,

    # margin inside the heart
    r_scaled = 0.99 * r_expanded,

    heart_radius = R_heart(theta),

    x_heart = r_scaled * heart_radius * cos(theta),
    y_heart = r_scaled * heart_radius * sin(theta)
  )
```


**rebuild community_anchors**

**add/subtract any communities as per count above**

```{r}
community_anchors <- tibble::tibble(
  community = 1:10,

  community_name = c(
    "Community 1",
    "Community 2",
    "Community 3",
    "Community 4",
    "Community 5",
    "Community 6",
    "Community 7",
    "Community 8",
    "Community 9",
    "Community 10"
  ),

  # Positions distributed through the heart
  anchor_x = c(
    -0.92,   # upper upper left
    -0.72,   # upper upper left
    -0.52,   # upper left
     0.52,   # upper right
     0.00,   # centre
    -0.34,   # lower left
     0.05,   # lower right
     0.07,   # lower lower right
     0.09,  # lower lower right
     0.10    # lowest lower right
  ),

  anchor_y = c(
     0.30,
     0.30, 
     0.30,
     0.30,
     0.02,
    -0.34,
    -1.05,
    -1.05,
    -1.05,
    -1.05
  )
)
```

**rebuild community_centers**

```{r}
names(layout_heartmap_full)
```

```{r}
layout_heartmap_full <- layout_heartmap_full |>
  mutate(
    community = as.integer(Group),
    community_name = paste("Community", community)
  )
```

```{r}
layout_heartmap_full |>
  distinct(community, community_name) |>
  arrange(community)
```

```{r}
community_centres <-
  layout_heartmap_full |>
  group_by(
    community,
    community_name
  ) |>
  summarise(
    centre_x = mean(x_heart, na.rm = TRUE),
    centre_y = mean(y_heart, na.rm = TRUE),
    .groups = "drop"
  )
```

```{r}
community_centres
```

## Communities 1 **Layout spread**

```{r}
layout_spread <-
  layout_heartmap_full |>
  left_join(
    community_centres,
    by = c(
      "community",
      "community_name"
    )
  ) |>
  left_join(
    community_anchors |>
      select(
        community,
        target_x = anchor_x,
        target_y = anchor_y
      ),
    by = "community"
  ) |>
  mutate(
    target_x = coalesce(target_x, centre_x),
    target_y = coalesce(target_y, centre_y),

    dx = x_heart - centre_x,
    dy = y_heart - centre_y,

    spread_factor =
      if_else(
        community == 5,
        1.55,
        1.25
      ),
    nudge_strength = 0.19,

    x_tmp = centre_x + spread_factor * dx,
    y_tmp = centre_y + spread_factor * dy,

# Nudge community centre toward target
    x_tmp =
      x_tmp +
      nudge_strength * (target_x - centre_x),

    y_tmp =
      y_tmp +
      nudge_strength * (target_y - centre_y),

# Expand the complete network outward
    outward_scale = 1.45,

    x_tmp = x_tmp * outward_scale,
    y_tmp = y_tmp * outward_scale,

# Constrain back inside heart
    theta_tmp = atan2(y_tmp, x_tmp),
    r_tmp = sqrt(x_tmp^2 + y_tmp^2),
    r_allowed = 0.91 * R_heart(theta_tmp),

    shrink = if_else(
      r_tmp > 0,
      pmin(1, r_allowed / r_tmp),
      1
    ),

    x_plot = x_tmp * shrink,
    y_plot = y_tmp * shrink
  )
```

```{r}
names(layout_spread)
```

## Rebuild edge coordinates for ggplot

```{r}
node_xy <- layout_spread |>
  select(
    name,
    community,
    Count,
    x_plot,
    y_plot
  )

edge_plot <- edges_full |>
  left_join(
    node_xy |>
      select(
        name,
        x_from = x_plot,
        y_from = y_plot
      ),
    by = c("from" = "name")
  ) |>
  left_join(
    node_xy |>
      select(
        name,
        x_to = x_plot,
        y_to = y_plot
      ),
    by = c("to" = "name")
  ) |>
  filter(
    !is.na(x_from),
    !is.na(y_from),
    !is.na(x_to),
    !is.na(y_to)
  )

```

## 5. Choose labels

```{r}
label_nodes <- layout_spread |>
  slice_max(Count, n = 100, with_ties = FALSE)
```

**name categories**
**add/subtract any communities as per count above**

```{r}
community_labels <- c(
    "Community 1",
    "Community 2",
    "Community 3",
    "Community 4",
    "Community 5",
    "Community 6",
    "Community 7",
    "Community 8",
    "Community 9",
    "Community 10"
  )
```

**make soft pretty heart**

```{r}
heart_nice <- heart_boundary |>
  mutate(
    x = 1.02 * x,
    y = 1.02 * y
  )
```

**add/subtract any communities as per count above**

```{r}
# Current community IDs
groups <- sort(unique(layout_spread$community))

#Nine colours
FBH_cols <- c(
  "Community 1" = "#7c0063",   # burgundy
  "Community 2" = "#e2aad7",   # pink
  "Community 3" = "#03d9f5",   # blue
  "Community 4" = "#ffbe0b",
  "Community 5" = "#fd8a09",
  "Community 6" = "#fb5607",
  "Community 7" = "#5f5ff6",
  "Community 8" = "#7dba60",
  "Community 9" = "#ff006e",
  "Community 10" = "#480ca8"
)

# Ensure the palette follows the communities in the data
group_cols <- FBH_cols[paste("Community", groups)]
names(group_cols) <- as.character(groups)
```

```{r}
group_cols
```


## 6. Pretty heart plot

```{r}
FBH.p <- ggplot() +
  geom_path(
    data = heart_nice,
    aes(x = x, y = y),
    colour = "#C7A6E8",
    linewidth = 1.3,
    alpha = 0.18,
    lineend = "round",
    linejoin = "round"
  ) +
  geom_path(
    data = heart_nice,
    aes(x = x, y = y),
    colour = "#A96BEA",
    linewidth = 0.65,
    alpha = 0.85,
    lineend = "round",
    linejoin = "round"
  ) +

# community background blobs
  
  ggforce::geom_mark_hull(
    data = layout_spread,
    aes(
      x = x_plot,
      y = y_plot,
      group = as.factor(community),
      fill = as.factor(community)
    ),
    concavity = 2,
    expand = grid::unit(11, "mm"),
    radius = grid::unit(16, "mm"),
    alpha = 0.035,
    colour = NA,
    show.legend = TRUE
  ) +
    # 2. internal gradient washes
  stat_density_2d(
    data = layout_spread |> filter(community == groups[1]),
    aes(
      x = x_plot,
      y = y_plot,
      alpha = after_stat(0.80 * nlevel)
    ),
    geom = "polygon",
    fill = group_cols[groups[1]],
    contour_var = "ndensity",
    bins = 5,
    colour = NA,
    show.legend = FALSE
  ) +

  stat_density_2d(
    data = layout_spread |> filter(community == groups[2]),
    aes(
      x = x_plot,
      y = y_plot,
      alpha = after_stat(1.00 * nlevel)
    ),
    geom = "polygon",
    fill = group_cols[as.character(groups[2])],
    contour_var = "ndensity",
    bins = 5,
    show.legend = FALSE
  ) +

  stat_density_2d(
    data = layout_spread |> filter(community == groups[3]),
    aes(
      x = x_plot,
      y = y_plot,
      alpha = after_stat(1.10 * nlevel)
    ),
    geom = "polygon",
    fill = group_cols[as.character(groups[3])],
    contour_var = "ndensity",
    bins = 5,
    show.legend = FALSE
  ) +
  stat_density_2d(
  data = layout_spread |>
    filter(community == groups[4]),
  aes(
    x = x_plot,
    y = y_plot,
    alpha = after_stat(1.00 * nlevel)
  ),
  geom = "polygon",
  fill = group_cols[as.character(groups[4])],
  contour_var = "ndensity",
  bins = 5,
  colour = NA,
  show.legend = FALSE
) +

stat_density_2d(
  data = layout_spread |>
    filter(community == groups[5]),
  aes(
    x = x_plot,
    y = y_plot,
    alpha = after_stat(1.00 * nlevel)
  ),
  geom = "polygon",
  fill = group_cols[as.character(groups[5])],
  contour_var = "ndensity",
  bins = 5,
  colour = NA,
  show.legend = FALSE
) +

  # edges
  geom_segment(
    data = edge_plot,
    aes(
      x = x_from,
      y = y_from,
      xend = x_to,
      yend = y_to,
      linewidth = weight
    ),
    colour = "grey35",
    alpha = 1.1,
    show.legend = FALSE
  ) +

# ----------------------------------------------------------
# Broad, soft community clouds
# ----------------------------------------------------------

ggforce::geom_mark_hull(
  data = layout_spread,
  aes(
    x = x_plot,
    y = y_plot,
    group = factor(community),
    fill = factor(community)
  ),
  concavity = 1.8,
  expand = grid::unit(19, "mm"),
  radius = grid::unit(25, "mm"),
  alpha = 0.025,
  colour = NA,
  show.legend = FALSE
) +

# A slightly tighter second cloud adds gentle depth
ggforce::geom_mark_hull(
  data = layout_spread,
  aes(
    x = x_plot,
    y = y_plot,
    group = factor(community),
    fill = factor(community)
  ),
  concavity = 2.2,
  expand = grid::unit(10, "mm"),
  radius = grid::unit(17, "mm"),
  alpha = 0.075,
  colour = NA,
  show.legend = FALSE
) +

  # nodes
  geom_point(
    data = layout_spread,
    aes(
      x = x_plot,
      y = y_plot,
      colour = as.factor(community),
      size = Count
    ),
    alpha = 0.88
  ) +

  # word labels
  geom_label_repel(
    data = label_nodes,
    aes(
      x = x_plot,
      y = y_plot,
      label = name,
      colour = as.factor(community)
    ),
    fill = "white",
    label.size = NA,
    alpha = 0.90,
    size = 5.1,
    max.overlaps = Inf,
    min.segment.length = 0,
    segment.alpha = 0.28,
    show.legend = FALSE
  ) +

  # heart outline drawn last so it stays crisp
  #geom_path(
  #  data = heart_boundary,
  #  aes(x = x, y = y),
  #  colour = "purple",
  #  linewidth = 0.7,
  #  alpha = 0.65
  #) +

  scale_colour_manual(
  values = group_cols,
  labels = community_labels,
  name = "Community"
) +
scale_fill_manual(
  values = group_cols,
  labels = community_labels,
  guide = "none"
) +
scale_alpha_continuous(
  range = c(0.012, 0.12),
  guide = "none"
) +
scale_size_continuous(
  range = c(1.5, 8),
  name = "Count"
) +
scale_linewidth_continuous(
  range = c(0.03, 0.45),
  guide = "none"
) +
guides(
  colour = guide_legend(
    override.aes = list(
      shape = 21,
      fill = unname(group_cols),
      colour = unname(group_cols),
      alpha = 0.35,
      size = 4,
      stroke = 0.6
    ),
    order = 2
  ),
  size = guide_legend(order = 1)
) +
labs(
  title = "A Heart-Shaped Keyword Network Monstrosity of Jane Austen Novels",
  subtitle = "Keywords are mapped into a heart-constrained network anlysis layout and grouped into ten thematic communities.\nNode size reflects word frequency; edges indicate co-occurrence relationships.",
  caption = "Communities identified using Louvain clustering. Figure created in R with a custom heart-constrained layout."
) +
coord_equal(clip = "off") +
theme_void(base_size = 12) +
theme(
  legend.position = "right",
  plot.background = element_rect(fill = "white", colour = NA),
  panel.background = element_rect(fill = "white", colour = NA),
  plot.margin = margin(20, 35, 20, 20),

  plot.title = element_text(
    size = 18,
    face = "bold",
    colour = "grey20",
    hjust = 0
  ),
  plot.subtitle = element_text(
    size = 11,
    colour = "grey35",
    hjust = 0,
    lineheight = 1.2
  ),
  plot.caption = element_text(
    size = 9,
    colour = "grey45",
    hjust = 0
  ),
  legend.title = element_text(
    size = 11,
    face = "bold"
  ),
  legend.text = element_text(
    size = 10
  )
)
```


## 7.Print

```{r}
ggsave(
  "C:/Users/u1130196/OneDrive - Australian National University/Documents/The Feedback Heart/FBH_Final.png",
  plot = FBH.p,
  width = 20,
  height = 14,
  dpi = 300
)
```


